---
title: "Thesis Analysis"
author: "Rory Butler"
date: "4/02/2022"
output: 
  html_document:
    keep_md: true
    toc: true
---
```{r global options, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(include=TRUE, warning=FALSE, message=FALSE, echo=FALSE)
```

## Pulling in all the data and transforming as needed
The food atlas data needs to be restricted to Colorado only
Shape file shouldn't need to be restricted

Poverty data we only need to keep poverty % value for each census tract
Education will need to transform into fewer categories and ultimately use just one split (college or not, probably)
Geographic mobility will transform into moved or didn't move
Race will transform into white % only (white vs non white essentially)

All the ACS files have a lead-in to the Census Tract # that differs from the atlas and obesity files

```{r Pull in all data}
library(tidyverse)
library(rgdal)
library(sf)
atlas <- read_csv('food_access_research_atlas.csv')
obesity_shape <- readOGR(".", "Obesity_in_Adults_-_CDPHE_Community_Level_Estimates_(Census_Tracts)")
race <- read_csv('acs2020_5yr_B03002_race.csv')
mobility <- read_csv('acs2020_5yr_B07003_geographic_mobility.csv')
education <- read_csv('acs2020_5yr_B15003_education.csv')
poverty <- read_csv('acs2020_5yr_B17001_poverty.csv')
```
```{r check data}
obesity_shape
```